---
bibliography: /home/persican/Documents/library.bib
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# eemR [![Travis-CI Build Status](https://travis-ci.org/PMassicotte/gtrendsR.svg?branch=master)](https://travis-ci.org/PMassicotte/eemR) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/PMassicotte/eemR?branch=master&svg=true)](https://ci.appveyor.com/project/PMassicotte/eemR) [![Package-License](https://img.shields.io/badge/license-GPL%20%28%3E=%202%29-brightgreen.svg?style=flat)](http://www.gnu.org/licenses/gpl-2.0.html)

The **eemR** package implements various functions used calculate metrics from excitation-emission matrix (EEM) as well as to preform pre-processing corrections before PARAFAC analysis [@Bro1997; @Stedmon2005; @Murphy2013]. All functions from this package start with the `eem_` prefix.

```{r}
library(eemR)
ls("package:eemR")
```

The package can be installed using the following command.

```{r install, eval = FALSE}
devtools::install_github("PMassicotte/eemR")
```

Please note this is a very alpha version of the package for testing purpose only.

# Reading EEMs

At the moment, the following EEM types are supported:

- Cary Eclipse `.csv` files

- Aqualog `.dat` files

- Shimadzu `.TXT` files

EEM can be read using the `eem_read()` function. *Please fill an [issue](https://github.com/PMassicotte/eemR/issues) if you have other file formats you would like to add to the package*.

At the moment I need files from:

- FluoromaxIII

- Perkin elmer

- Hitachi

```{r}

library(eemR)

## Reading a single eem
file <- system.file("extdata/cary/eem", "sample1.csv", package = "eemR")
eem <- eem_read(file)

plot(eem)

## Reading a folder
folder <- system.file("extdata/cary/eem", package = "eemR")
eem <- eem_read(folder)

plot(eem) ## Plot the first eem
plot(eem, which = 2) ## Plot the second eem

## Aqualog EEM
folder <- system.file("extdata/aqualog", package = "eemR")
eem <- eem_read(folder)

plot(eem) ## Plot the first eem

```

# Shimadzu files

Some spectrophotometers (such as Shimadzu) do not include excitation wavelengths in the fluorescence files. In these cases, we can use the `eem_set_wavelengths()` function to manually provide vectors of emission and/or excitation wavelengths.

```{r}
folder <- system.file("extdata/shimadzu", package = "eemR")
eems <- eem_read(folder)

eems <- eem_set_wavelengths(eems, ex = seq(230, 450, by = 5))
```

# Manipulating EEMs

Extracting of removing EEMs can be useful when reading a bunch of files containing both measurements and blank fluorescence. This can be done easily using `eem_extract()`. For example, lets read a whole folder and then remove the blank water.

```{r}
folder <- system.file("extdata/cary", package = "eemR")

eems <- eem_read(folder, recursive = TRUE)

blank <- eem_extract(eems, "nano", remove = FALSE)

eems <- eem_extract(eems, "nano", remove = TRUE)


## Remove sample 1 to 3
res <- eem_extract(eems, 1:3, remove = TRUE)
```

# Implemented metrics

The current implemented metrics are:

1. The fluorescence index (FI) developed by @McKnight2001.

2. The fluorescence peaks proposed by @Coble1996.

3. The fluorescence humification index (HIX) by @Ohno2002.

4. The biological fluorescence index (BIX) by @Huguet2009.

```{r, warning = FALSE}

library(eemR)

folder <- system.file("extdata/cary/eem", package = "eemR")
eem <- eem_read(folder)

eem_fluorescence_index(eem, verbose = FALSE)

eem_coble_peaks(eem, verbose = FALSE)

eem_humification_index(eem, verbose = FALSE)

eem_humification_index(eem, verbose = FALSE, scale = TRUE)

eem_biological_index(eem, verbose = FALSE)

```

# PARAFAC pre-processing

Three types of correction are currently supported:

1. `eem_remove_blank()` which subtract a water blank from the eem.

2. `eem_remove_scattering()` which remove both *Raman* and *Rayleigh* scattering.

3. `eem_raman_normalisation()` which normalize EEM fluorescence intensities [@Lawaetz2009].

4. `eem_inner_filter()` which correct for both primary and secondary inner-filter effect.

## Blank removal

The `eem_remove_blank()` function subtract blank (miliq) water from eem. Scatter bands can often be reduced by subtracting water blank [@Murphy2013].

```{r, out.width='300cm', out.height='250cm', fig.show='hold'}
file <- system.file("extdata/cary/eem", "sample1.csv", package = "eemR")
eem <- eem_read(file)

file <- system.file("extdata/cary", "nano.csv", package = "eemR")
blank <- eem_read(file)

res <- eem_remove_blank(eem, blank)

plot(eem)
plot(res)
```

## Removing Raman and Rayleigh scattering (1st and 2nd order)

The `eem_remove_scattering()` function removes both Raman and Rayleigh scattering from EEMs.

```{r}

res <- eem_remove_scattering(eem = eem, type = "raman", order = 1, width = 10)
res <- eem_remove_scattering(eem = res, type = "rayleigh", order = 1, width = 10)

plot(res)
```

## Raman normalization

The `eem_raman_normalisation()` function implement a simple calibration method for fluorescence intensity using only the integrated area of a water Raman peak. More details can be found in @Lawaetz2009.

```{r}

res <- eem_raman_normalisation(res, blank)

plot(res)

```

## Inner-filter effect correction

To account for reabsorption of the light emitted by fluorophores in the water, absorbance spectra are used for correction of both primary and secondary inner filtering effects in the EEMs [@Ohno2002; @Parker1957; @Kothawala2013].

```{r}
data("absorbance")

res <- eem_inner_filter_effect(eem = res,
                               absorbance = absorbance,
                               pathlength = 1) ## 1 cm fluo pathlenght

plot(res)

```

### Name matching

The names of `absorbance` variables are expected to match those of the eems. If the appropriate absorbance spectrum is not found, an uncorrected eem will be returned and a warning message will be printed.

### Sample dilution

@Kothawala2013 have shown that a 2-fold dilution was required for samples presenting total absorbance > 1.5. Accordingly, a message will warn the user if total absorbance is greater than this threshold.

## Export to Matlab

PARAFAC analysis was made easy with the fantastic Matlab [`drEEM`](http://www.models.life.ku.dk/drEEM) toolbox [@Murphy2013]. The function `eem_export_matlab()` can be  used to export the EEMs into a `m-file` directly usable in Matlab by the `drEEM` toolbox.

```{r}
folder <- system.file("extdata/cary/eem", package = "eemR")
eem <- eem_read(folder)

filename <- paste(tempfile(), ".mat", sep = "")

eem_export_matlab(filename, eem)

## It is also possible to export more than one object at time
eem_export_matlab(filename, eem, eem)
```

Note that the name of the structure generated by the function will be `OriginalData` to *complement* with PARAFAC standard. Then, the importation into Matlab is made easy using the `load()` function. Please note that there is a bug preventing to keep matrix dimension. Simply use the `reshape()` function after you exported data.

```matlab
load('FileName.mat');

OriginalData.X = reshape(OriginalData.X, ...
    OriginalData.nSample, ...
    OriginalData.nEm, ...
    OriginalData.nEx);

% Start PARAFAC analysis here...
```

# References
